---
title: "Data_prep"
author: "Dijia Tang"
date: "11/10/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
Load packeges
```{r}
library(zipcode)
library(tidyverse)
library(lubridate)
```
# Part 1

## Preliminary processing
Import food data and change the column names
```{r}
raw <- read.csv('https://s3.amazonaws.com/notredame.analytics.data/inspections.csv')
food <- raw
colnames(food) <- c('ID','DBAName','AKAName','License','FacilityType',
                    'Risk','Address','City','State','Zip',
                    'InspectionDate','InspectionType','Results','Violations',
                    'Latitude','Longitude','Location')
```
Change the data type for each column
```{r}
food <- food %>% 
  mutate(ID = as.integer(ID), DBAName = as.character(DBAName), AKAName = as.character(AKAName),
         License = as.character(License),FacilityType = as.factor(FacilityType), Risk = as.factor(Risk),
         Address = as.character(Address), City = as.factor(City), State = as.factor(State),
         Zip = as.factor(Zip), InspectionType = as.factor(InspectionType),
         Results = as.factor(Results), Violations = as.character(Violations), Latitude = as.double(Latitude),
         Longitude = as.double(Longitude), Location = as.character(Location))
food$InspectionDate = mdy(food$InspectionDate)
food <- food %>% 
  mutate(InspectionDate = as.Date(InspectionDate))
```
Output summary statistics
```{r}
summary(food)
```
# Part 2

Change city names to their title forms
```{r}
food <- food %>% 
  mutate(City = str_to_title(City))
table(food$City)
```
Replace incorrect spellings of 'Chicago'
```{r}
food$City <- str_replace_all(food$City,'Cchicago|Chchicago|Chcicago|Chicagoi|Chicagochicago', "Chicago")
```
Change the city name from Brideview to Bridgeview and add the Zip code
```{r}
food[which(grepl('Brideview', food$City)), 'City'] <- 'Bridgeview'
food <- food %>% 
  mutate(Zip = as.character(Zip))
food[which(grepl('Bridgeview', food$City)), 'Zip'] <- '60455'
```
Change Zip code 60627 to 60827
```{r}
food <- food %>% 
  mutate(Zip = str_replace_all(Zip, '60627', '60827'))
food <- food %>% 
  mutate(Zip = as.factor(Zip))
```
Create a dataframe that contains missing values for City
```{r}
MissingCity <- food[which(is.na(food$City)),]
```
All missing cities have zipcodes. Use zipcode dataset to match those cities.
```{r}
data("zipcode")
cityjoin <- merge(x = MissingCity, y = zipcode, by.x='Zip', by.y = 'zip')
cityjoin <- cityjoin %>% 
  select(-City,-latitude,-longitude,-state) 
cityjoin = cityjoin[c(2,3,4,5,6,7,8,17,9,1,10,11,12,13,14,15,16)]
names(cityjoin)[names(cityjoin) == 'city'] <- 'City'
food <- food %>% 
  filter(!is.na(City)) %>% 
  rbind(cityjoin)
```
Examine missing cities
```{r}
table(is.na(food$City))
```
Create a dataframe that contains missing values for State
```{r}
MissingState <- food[which(is.na(food$State)),]
```
We can see that since the city is Chicago for all rows in MissingState, we should assign 'IL' to the State variable. Find the state name using indexation.
```{r}
food[which(is.na(food$State)),'State'] <- as.character(na.omit(unique(food[which(grepl(unique(MissingState$City), food$City)), 'State'])))
table(is.na(food$State))
```
Create a dataframe that contains missing values for Zip
```{r}
MissingZip <- food[which(is.na(food$Zip)),]
```
There are two restaurants in MissingZip. Their licenses are '1166638' and '1515116'.

**Strategy 1:** Let's explore if they appear elsewhere.
```{r}
store_1166638 <- food %>% 
  filter(License == '1166638')
store_1515116 <- food %>% 
  filter(License == '1515116') 
```
We can't find these records elsewhere in the original dataset.

**Strategy 2:** Try to match based on latitude and longitude in zipcode.
```{r}
ziptest <- zipcode %>% 
  mutate(longitude = round(as.double(longitude),2),
         latitude = round(as.double(latitude),2))
zipselect <- ziptest %>% 
  filter(longitude == -87.67 & latitude == 42.02)
```
We can't find the zipcodes based on latitude/longitude matching.

**Strategy 3:** Find the Zipcode for MissingZip based on the closest house number and street name (hot-deck).
```{r}
m1 <- unique(food[grep('755\\d\\sN\\sPAULINA\\sST',food$Address),]$Zip)
m2 <- food[grep('..\\d\\d\\sN\\sFREMONT\\sST',food$Address),]$Zip 
m3 <- m2 %>% 
  unique() %>% 
  na.omit()
```
Assign the above zipcodes to the two licenses.
```{r}
food[which(grepl('1515116', food$License)), 'Zip'] = m1 
food[which(grepl('1166638', food$License)), 'Zip'] = m3
```
Examine missing zipcodes
```{r}
table(which(is.na(food$Zip)))
```

Create a dataframe that contains missing values for Longitude
```{r}
MissingLong <- food[which(is.na(food$Longitude)),]
```
All MissingLongs have zipcodes and have neither latitude nor location. Find latitude, longitude, and location based on zip.
```{r}
longjoin <- merge(x = MissingLong, y = zipcode, by.x='Zip', by.y = 'zip')
longjoin <- longjoin %>% 
  select(-City,-Latitude,-Longitude,-state) 
longjoin = longjoin[c(2,3,4,5,6,7,8,15,9,1,10,11,12,13,16,17,14)]
names(longjoin)[names(longjoin) == 'city'] <- 'City'
names(longjoin)[names(longjoin) == 'latitude'] <- 'Latitude'
names(longjoin)[names(longjoin) == 'longitude'] <- 'Longitude'
longjoin <- longjoin %>% 
  mutate(Location = as.character((paste("(",Latitude,',',Longitude,")",sep = ''))))
glimpse(longjoin)
food <- food %>% 
  filter(!is.na(Longitude)) %>% 
  rbind(longjoin)
table(is.na(food$Longitude))
```
Explore missing values for Latitude - 0 records
```{r}
food[which(is.na(food$Latitude)),]
```
Explore missing values for Longitude - 0 records
```{r}
food[which(is.na(food$Longitude)),]
```
Explore missing values for Location - 0 records
```{r}
food[which(is.na(food$Location)),]
```
## Output lists for unqiue values in City, State, and Zip, sorted alphabetically
```{r}
City <- as.list(unique(as.character(sort(food$City))))
State <- as.list(unique(as.character(sort(food$State))))
Zip <- as.list(unique(as.character(sort(food$Zip))))
City
State
Zip
```
## Output ranges of latitude and longitude
```{r}
range(food$Latitude)
range(food$Longitude)
```
# Part III

Add ViolationCount and create a boxplot
```{r}
food <- food %>% 
  rowwise() %>% 
  mutate(ViolationCount = str_count(Violations,pattern = '\\|')+1) 
food$ViolationCount <- replace_na(food$ViolationCount,0)

plotdata <- food %>% 
  select(Results, ViolationCount) %>% 
  filter(Results == 'Pass' | Results == 'Fail' | Results == 'Pass w/ Conditions') 

ggplot(plotdata)+
  geom_boxplot(aes(x = Results, y = ViolationCount))
```
**Discussion on the boxplot**

We can see that from a overall perspective, Pass corresponds to the fewest ViolationCount, while Fail corresponds to the highest. Pass w/ Conditions are somewhere in between.

We can tell that whether the restaurant passes the inspection really depends on the number of violations it has.

The fewer the violations, the more likely the restaurant is going to pass the inspection.

## Final adjustments and examination of food
```{r}
food <- food %>% 
  mutate(ViolationCount = as.integer(ViolationCount))
summary(food)
str(food)
```
